---
title: "kuenm: modeling process"
output:
  html_document:
      toc: yes
      toc_depth: 4
---

The R markdown file is created in the working directory, and is designed to make the processes of model calibration and final model creation more reproducible.

Information on using this R Markdown file:

- Try executing code chunks by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*.
- Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

A brief tutorial for using functions of the kuenm R package can be found in the <a href="https://github.com/marlonecobos/kuenm" target="_blank">package vignette</a>. Additionally, function help can be checked to change arguments according to specific needs.

<br>

### Candidate models

Candidate models are a large set of candidate models created to respond to the need to test broad suites of parameter combinations, such as, distinct regularization multiplier values, various feature classes, and different sets of environmental variables. The following code calls the help page of the function kuenm_cal.

```{r, eval=FALSE}
help(kuenm_cal)
```

<br>

The next chunk of code is for preparing the arguments for using the function following the modularity principle. These variables can be changed according to each case.

```{r, eval=FALSE}
occ_joint <- "cmex_joint.csv"
occ_tra <- "cmex_train.csv"
M_var_dir <- "M_variables"
batch_cal <- "Candidate_models"
out_dir <- "Candidate_Models"
reg_mult <- c(seq(0.1, 1, 0.3), seq(2, 4, 1))
f_clas <- "no.t.h"
background <- 10000
maxent_path <- "C:/Maxent/3.4.1/"
wait <- FALSE
run <- TRUE
```

<br>

The following is the code for using the function.

```{r, eval=FALSE}
kuenm_cal(occ.joint = occ_joint, occ.tra = occ_tra, M.var.dir = M_var_dir, batch = batch_cal,
          out.dir = out_dir, reg.mult = reg_mult, f.clas = f_clas, background = background,
          maxent.path = maxent_path, wait = wait, run = run)
```

<br>

### Evaluation and selection of best models

Evaluation is a crucial step in model calibration. This step centers on selecting candidate models and their associated parameters to identify the best models for the purposes of the study. The kuenm_ceval function evaluates candidate models based on three distinct criteria: statistical significance (based on partial ROC analyses), prediction ability (we use omission rates, but other metrics, such as overall correct classification rate, can also be used), and model complexity (here evaluated using AICc). The following code chunk calls the function help window.

```{r, eval=FALSE}
help(kuenm_ceval)
```

<br>

Below, arguments for this functions will be defined.

```{r, eval=FALSE}
occ_test <- "cmex_test.csv"
out_eval <- "Calibration_results"
threshold <- 5
rand_percent <- 50
iterations <- 500
kept <- TRUE
selection <- "OR_AICc"
paral_proc <- FALSE # make this true to perform MOP calculations in parallel, recommended
                    # only if a powerfull computer is used (see function's help)
# Note, some of the variables used here as arguments were already created for the previous function
```

<br>

This code also allows evaluating candidate models that were created previously, selecting those with best performance based on the three criteria.

```{r, eval=FALSE}
cal_eval <- kuenm_ceval(path = out_dir, occ.joint = occ_joint, occ.tra = occ_tra, occ.test = occ_test,
                        batch = batch_cal, out.eval = out_eval, threshold = threshold,
                        rand.percent = rand_percent, iterations = iterations, kept = kept,
                        selection = selection, parallel.proc = paral_proc)
```
